Quotient Normalized Maximum Likelihood Criterion for Learning Bayesian Network Structures

### 2017-03-16 qnml.tex
consistency proof added

\subsection {On implementation}

>>>>

One important property possessed by nearly every model selection criterion is consistency. In our context, consistency means that given a data matrix with $N$ samples coming from a distribution faithful to some DAG $G$, the qNML will give the highest score to true graph $G$ with a probability tending to one as $N$ increases. We will show this by first proving that qNML is asymptotically equivalent to widely used BIC criterion which is known to be consistent \cite{Schw78, Haug88}.

The BIC can be written as
\begin{equation}
\textnormal{BIC}(D,G) = \sum_{i = 1}^n \log P(D_i \ | \ \hat{\theta}_{i | G_i} ) - \frac{q_i(r_i - 1)}{2} \log N,
\end{equation}where $D_i = D[\cdot, i]$ stands for the data on variable $i$, $\hat{\theta}_{i | G_i}$ denotes the maximum likelihood parameters of the conditional distribution of variable $i$ given its parents in $G$. Since the both BIC and qNML scores are decomposable, we can focus on studying the local scores. Using the definition of qNML, we have
\begin{align}\label{qnmlDef2}
s^{qNML}_i(D;G) &= \log \frac{P(D_{i, G_i} \ | \ \hat{\theta}_{i, G_i} )}{P(D_{G_i} \ | \ \hat{\theta}_{G_i} )} \notag \\
& -(reg(N,q_i r_i) - reg(N,q_i)).
\end{align}

Now, we observe that the first term in (\ref{qnmlDef2}) can written as
$$
\log \frac{P(D_{i, G_i} \ | \ \hat{\theta}_{i, G_i} )}{P(D_{G_i} \ | \ \hat{\theta}_{G_i} )} = \log P(D_i \ | \ \hat{\theta}_{i | G_i} )
$$ which simply shows that the goodness-of-the-fit terms agree between the criteria. Next, we consider the difference of regrets in (\ref{qnmlDef2}) which corresponds to the penalty term of BIC. The following theorem states that these two are asymptotically equal.

\begin{theorem}
Let $r_i$ and $q_i$ denote the number of possible values for variable $X_i$ and the possible configurations of parents $X_{G_i}$, respectively. As $N \to \infty$,
$$reg(N,q_i r_i) - reg(N,q_i) = \frac{q_i(r_i - 1)}{2}\log N + O(1).$$
\end{theorem}   
\begin{proof}
By considering only the leading term in (\ref{eqn:szp}), the regret for a single multinomial variable with $m$ categories can be written asymptotically as
$$
reg(N,m) = \frac{m-1}{2}\log N + O(1).
$$ Using this, we have
\begin{align*}
reg(N,q_i r_i)& -  reg(N,q_i) \\ 
&= \frac{q_ir_i-1}{2}\log N-\frac{q_i-1}{2}\log N + O(1) \\
&= \frac{q_ir_i-1-q_i + 1}{2}\log N + O(1) \\
&= \frac{q_i(r_i - 1)}{2} \log N + O(1).
\end{align*}
\end{proof}
This suffices in order to show that qNML is asymptotically equivalent to BIC and, hence, consistent. The outline of this proof follows the similar pattern as found in \cite{SilanderIJAR10} where the consistency of fNML was proved.

\subsection{On implementation}
